---
title: "My Course Template, Session 2"
subtitle: " <html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html> Bioinformatics Resource Center - Rockefeller University"
author: 
  - "http://rockefelleruniversity.github.io/RU_course_template/"
  - "brc@rockefeller.edu"
output: 
  xaringan::moon_reader:
    css: ["default.css", "metropolisCustom.css", "metropolis-fontsCustom.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      titleSlideClass: [middle, inverse]
  html_document:
    toc: true # table of content true
    toc_float: yes
    depth: 3  # upto three depths of headings (specified by #, ## and ###)
    number_sections: false  ## if you want number sections at each table header
    theme: united  # many options for theme, this one is my favorite.
    highlight: tango  # specifies the syntax highlighting style
params:
  isSlides: "no"
---

```{r setup, include=FALSE}
suppressPackageStartupMessages(require(knitr))
knitr::opts_chunk$set(echo = TRUE, tidy = T)
```

```{r, results='asis',include=TRUE,echo=FALSE}
if(params$isSlides != "yes"){
  cat("# ?? name of course?? (part 2)

---
"    
  )
  
}

```


##  Recap

?? can add a recap to link to keypoints in first session??
Session 1 covered introduction to R data types and import/export of data.

- [Background to R](https://rockefelleruniversity.github.io/Intro_To_R_1Day/r_course/presentations/singlepage/introToR_Session1.html#background-to-r)
- [Data types in R](https://rockefelleruniversity.github.io/Intro_To_R_1Day/r_course/presentations/singlepage/introToR_Session1.html#data_types_in_r)
- [Reading and writing in R](https://rockefelleruniversity.github.io/Intro_To_R_1Day/r_course/presentations/singlepage/introToR_Session1.html#reading-and-writing-data-in-r)


---
## Same stuff
??More content for second session??

---
## Time for an exercise!


Exercise on functions can be found [here](../..//exercises/exercises/MyExercise3_exercise.html)

---
## Answers to exercise.


Answers can be found here  [here](../..//exercises/answers/MyExercise3_answers.html)

Rcode for answers can be found here  [here](../..//exercises/answers//answers/MyExercise3_Answers.R)

